---
title: "R packages: Part I"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide

```{r child = "../setup.Rmd"}
```

```{r echo = FALSE, message = FALSE}
library(tidyverse)
options(scipen = 999)
```

<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

class: middle

## Naming your package

```{r eval=FALSE}
available::available("mypackage")
```

---

class: middle

## Creating your package

File > New Project > New Directory > R Package

---

class: middle

## Package directory

- `R` where R scripts for functions are stored. 
- `man` where documentation files are stored.
- DESCRIPTION contains the metadata about the package. 
- NAMESPACE will be generated automatically. More on this later.

More components to follow later (e.g. data storage).


---

class: middle

## Shortcuts

Install Package: `Cmd + Shift + B`
Check Package:  `Cmd + Shift + E`
Test Package: `Cmd + Shift + T`

---

class: middle

## Adding a license

```{r eval=FALSE}
usethis::use_mit_license("Mine Dogucu")
```


```{r eval=FALSE}
✓ Writing 'LICENSE' # what CRAN requires
✓ Writing 'LICENSE.md' # created only for human-readers
✓ Adding '^LICENSE\\.md$' to '.Rbuildignore' # LICENSE.md is ignored when package is built
```

[usethis license functions](https://usethis.r-lib.org/reference/licenses.html)

---

class: middle

## Writing functions

- Functions are written in an `.R` file and stored in the `R` file.

- You do not need one `.R` file per function. If functions are closely related you may want to keep them in a single file. 

---


class: middle

## Documenting functions

When your cursor is inside a function code, then click on `Code>Insert ROxygen Skeleton`


---
class: middle

`#' Mean and Median Summaries for Numerical Vectors`  
`#'`  
`#' @param some_variable a numeric vector`  
`#' @param na.rm indicates whether missing data should be removed or not`  
`#'`
`#' @return a numeric vector with mean and median values`  
`#' @export`  
`#' @importFrom stats median dbinom`  
`#' @examples`  
`#' print_mean_median(mtcars$mpg, na.rm = TRUE)`

---

class: middle

## Dependencies

- All functions that are from packages other than `base` need to be explicitly imported, either as a function or as a full package.

- All functions from a package can be imported with `@import`

- Specific functions can be imported with `@importFrom stats median dbinom`

---

class: middle

## Documentation

```{r eval=FALSE}
devtools::document()
```

---

class: middle

## Workflow so far

1. Write function
2. Add ROxygen Skeleton and write documentation
3. `devtools:document()`
4. `devtools:check()`

---

class: middle

## Testing

```{r eval=FALSE}
devtools::test()
x No testing infrastructure found. Create it?

1: Yes
2: No
```

This will create the `tests` folder where we will store the previously written tests.
We will use the tests from last week.

---

class: middle

## Workflow 

1. Write function
2. Add ROxygen Skeleton and write documentation
3. `devtools:document()`
4. `devtools:check()`
5. Write tests
6. `devtools:test()`